---
title: "Function Showcase"
author: "Programmer C"
date: "2023-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Function Introduction
We use the built-in dataset 'economics' as our analysis basis, which was originally produced from US economic time series. The function here is to filter out the highest per capita consumption data for a given year with the first input argument 'y1', and display the month-on-month change of unemployment rate with the second input argument 'y2'.

In the following case, y1 is given the value '1980' and y2 is given the value '2012'. The final output consists of a list of two data tables, the first of which shows the highest per capita consumption value in the US in 1980, and the second of which shows the monthly change in the broad unemployment rate in 2012.

```{r showcase, message = FALSE}
#Load the dplyr package.
library(dplyr)
#Load the ggplot2 package.
library(ggplot2)
#Load the lubridate package.
library(lubridate)

#Load the dataset "economics".
data("economics")

#Define a function that filters for the highest per capita consumption value in a given year.
#Add a variable y2 to make it output the month-on-month change in the broad unemployment rate for the selected year.
transformed_economics <- function(y1, y2){
  #Transform 1.
  transformed_data <- economics %>% 
    #Add a new colunm to display per capita consumption.
    mutate(per_capita_consumption = economics$pce*10^9/(economics$pop*1000)) %>%
    #Filter the row with the highest per capita consumption.
    filter(year(as.Date(date)) == y1) %>% 
    filter(per_capita_consumption == max(per_capita_consumption)) %>%
    #Show the final tidy chart using 'summarise()' function. 
    summarise(Year = y1, 
              Date = date, 
              Max_pcc = per_capita_consumption)
  #Transform 2.
  transformed_data2 <- economics %>%
    #Add a column showing the month-on-month change in the broad unemployment rate.
    mutate(unemployment_change = (economics$unemploy/economics$pop)*100 - lag(economics$unemploy/economics$pop)*100) %>%
    #Filter out all data for the selected year. 
    filter(year(date) == y2) %>%
    #Remove the first month's data as the month-on-month change has no comparator in the first month.
    filter(!is.na(unemployment_change)) %>%
    group_by(date) %>%
    #Show the final tidy chart.
    summarise(unemployment_change)
  result <- list(transformed_data, transformed_data2)
  return(result)
}

#Preview the first few rows of transformed data.
print(transformed_economics(1980, 2012))
```

You can change the year variables in the bracket to get different output. For example, change the input variables to 1975 and 2013.

```{r showcase2, message=FALSE}

#Preview the first few rows of transformed data with different input.
print(transformed_economics(1975, 2013))

```

## Appendix: All code in this showcase.

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE} 
# this chunk generates the complete code appendix. 
```